---
title: "working_reimbursement_data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

install.packages("tidycensus")

```

# Load in the required packages to clean the data

```{r}
library(tidyverse)
library(vembedr)
library(janitor)
library(dplyr)
library(tidycensus)

census_api_key("549950d36c22ff16455fe196bbbd01d63cfbe6cf")

```

# Read in the dirty raw data

```{r}
# Read in dirty dataset
tidied_reimbursement_requests_redacted <- read_csv("data/tidied_reimbursement_requests_redacted.csv")

tidied_reimbursement_requests_redacted

```

# Glimpse the data

```{r}
glimpse(tidied_reimbursement_requests_redacted)


# We see that it has read in the time (in days) and all of the money values as characters, we want it to be viewed as a number.

```
# Clean up the data

```{r}

clean_reimbursement_requests <- tidied_reimbursement_requests_redacted %>%
  # clean up those unsightly column names using janitor function clean_names() to get rid of spaces & uppercase
  clean_names() %>%
  
  # get rid of empty rows using janitor function remove_empty()
  remove_empty() %>%
  
  # remove subtotals and fy rows
  filter(!is.na(location_of_incident)) %>%
  
  # remove $'s and commas for amount_of_request
  mutate(amount_of_request = str_remove_all(amount_of_request,"\\$|,|[a-z]|[A-Z]")) %>%
  mutate(amount_of_request = str_trim(amount_of_request)) %>%
  mutate(amount_of_request = na_if(amount_of_request,"")) %>%
  
  # remove $'s and commas for amount_approved
  mutate(amount_approved = str_remove_all(amount_approved,"\\$|,|[a-z]|[A-Z]")) %>%
  mutate(amount_approved = str_trim(amount_approved)) %>%
  mutate(amount_approved = na_if(amount_approved,"")) %>%

  # change to chr column
  mutate(amount_of_request = as.numeric(amount_of_request)) %>%
  mutate(amount_approved = as.numeric(amount_approved)) %>%
  
  # create new column where we collapse into just approved or denied without the extra addendums
  mutate(request_determination_cleaned = case_when(
    str_detect(request_determination,"Approved")~"Approved",
    str_detect(request_determination, "Denied")~"Denied"
  )) %>%
  
  mutate(zip_code = str_sub(location_of_incident, start = -5L, end = -1L)) %>%
  mutate(zip_code = case_when(
    str_detect(zip_code,"East") ~ "21228", TRUE ~ zip_code
  ))
  
  

```
# Get the number of approved and denied requests: 19 approved, 101 denied

```{r}

requests_approved_denied <- clean_reimbursement_requests %>%
  group_by(request_determination_cleaned) %>%
  count()


```

# Get the total amount paid out to approvals ever: $19,559.41

```{r}

total_amnt_paid <- clean_reimbursement_requests %>%
  select(amount_approved) %>%
  summarise(amnt_paid_in_total = sum(amount_approved, na.rm = TRUE))


```

# Get the avg amount requested from the city: ~$2,355.58

```{r}

avg_amnt_requested <- clean_reimbursement_requests %>%
  select(amount_of_request) %>%
  # divide the total amnt of money requested by total # of requests (120)
  summarise(avg_requested = sum(amount_of_request, na.rm = TRUE)/120)

```

# Get the avg amount paid per approval: ~$1029.44

```{r}

avg_paid_per_approval <- clean_reimbursement_requests %>%
  select(amount_approved) %>%
  #divide total amount approved by total # of approvals (19)
  summarise(avg_approved = sum(amount_approved, na.rm = TRUE)/19)
  
```

# Calculate zip code totals
```{r}

zip_code_totals <- clean_reimbursement_requests %>%
  group_by(zip_code, request_determination_cleaned) %>%
  count() %>%
  pivot_wider(names_from = request_determination_cleaned, values_from = n) %>%
  clean_names() %>%
  mutate(denied = replace_na(denied, 0), approved = replace_na(approved, 0)) %>%
  mutate(totals = approved + denied) %>%
  mutate(pct_approved = approved/totals*100)


```


# Using tidycensus, load in ZCTA API
```{r}

zcta_median_household_income <- get_acs(geography = "zcta",
              variables="B19013_001", year=2018, geometry = FALSE )

working_zcta_median_household_income <- zcta_median_household_income %>%
  clean_names() %>%
  mutate(name = str_remove(name, "ZCTA5 ")) %>%
  rename(median_household_income = estimate) %>%
  select(name, median_household_income)

```



```{r}

acs_variables <- load_variables(2018, "acs5")


```


```{r}

zcta_black_population <- get_acs(geography = "zcta", summary_var = "B02001_001",
              variables="B02001_003", year=2018, geometry = FALSE )

working_zcta_black_population <- zcta_black_population %>%
  clean_names() %>%
  mutate(name = str_remove(name, "ZCTA5 ")) %>%
  mutate(pct_black = estimate/summary_est*100) %>%
  select(name,pct_black)

```

# Left joining ZCTA and Baltimore zips
```{r}

joined_zip_codes <- zip_code_totals %>%
  left_join(working_zcta_median_household_income, by = c("zip_code" = "name")) %>%
  left_join(working_zcta_black_population, by = c("zip_code" = "name"))

```